---
title: "Supplement for Planting the seed for sound change: evidence from real-time MRI of velum kinematics in German"
output: 
  html_document: 
    highlight: tango
    number_sections: yes
    theme: paper
  pdf_document: 
    latex_engine: xelatex
    number_sections: yes
    toc: yes
bibliography: bib.bib
---

```{r setup, include=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
library(patchwork)
library(brms)
library(tidybayes)
library(extraDistr)
library(HDInterval)

core.num <- parallel::detectCores()
options(mc.cores = core.num)

my.seed <- 123
set.seed(my.seed)
```

# Prepare data

```{r read-data}
matdat <- read.csv("./rtMRI-velum/velum_data.csv", header = T)

# get rid of items that gesture onsets that begin *after* the vowel offset (only 5/7152 total items)
matdat <- matdat[(matdat$velumopening_gesture_on - matdat$Vokal_off) < 0,]

# separate coda contexts by alveolar voiced stop vs. alveolar voiceless stop
voiceless <- c("nt__", "nt_@", "nt_6", "nt_a")
voiced    <- c("nd_@", "nd_6", "nd_a")

matdat$voicing <- c()
matdat$voicing[matdat$post %in% voiceless]  <- "voiceless"
matdat$voicing[matdat$post %in% voiced]     <- "voiced"

# only include alveolar nasal items preceding a voiced or voiceless stop consonant
coda <- c(voiceless, voiced)

# only include neutrally stressed utterances
stresses <- c("N")

# subset the data
subdat <- matdat[matdat$post %in% coda & matdat$stress %in% stresses, ]
```

# Velum gesture duration

Since no specific information on expectations for velum gesture duration can be found in the literature, we followed more general expectations for speech segment durations, coded as very weakly informative priors.
The BRM for velum gesture duration was built using a log-normal distribution, since speech segment duration has been shown to be log-normally distributed [@Rosen2005; @Ratnikova2017; @Gahl2019].
The following distributions were used as weakly informative priors (on the log-odds scale): for the intercept of duration (corresponding to ND), a normal distribution with mean = 0 and standard deviation = 3 ($Normal(0, 3)$); for the effect of voicing (when NT), $Normal(0, 1)$.
These roughly correspond to a belief that the intercept of velum gesture duration (when the context is ND) is between 0 and 400 ms ($e^{3 \times 2} = 403$), and the duration changes (increases or decreases) by a factor of 1 to 7.4 in the ND context ($e^{1 \times 2} = 7.4$), at 95% confidence.
For the model standard deviation and the random intercept standard deviation we used a half-Cauchy distribution with location 0 and scale 0.1 ($HalfCauchy(0, 0.1)$); this corresponds to a 95% HDI = [0, 2.55] in log-odds, i.e. a factor change of 1 to 12.75 ($e^2.55 = 12.75$).

NOTE: Warnings about $\hat{R}$ and sample size are false warnings, and are not signs of problems with fitting and they can safely be ignored. See <https://github.com/paul-buerkner/brms/issues/864>.

```{r half-cauchy}
inverseCDF(c(0.025, 0.975), phcauchy, sigma = 0.1)
```

For the correlation between random effects, an $LKJ(2)$ distribution, as recommended by [@vasishth2018].
The same prior specification was used for the models of velum gesture onset time and velum gesture offset time.

```{r dur, cache=TRUE}
# create the dependent variable
subdat$DV <- subdat$velumopening_gesture_dur*1000

# full model
dur <- brms::brm(
  DV ~ voicing +
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 3), class = Intercept),
            prior(normal(0, 1), class = b, coef = voicingvoiceless),
            prior(cauchy(0, 0.1), class = sd),
            prior(cauchy(0, 0.1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 6000,
  warmup = 3000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/dur",
  save_all_pars = TRUE
)
```

```{r dur-summary}
dur
```

```{r dur-null, cache=TRUE}
# reduced/null model
dur_null <- brms::brm(
  DV ~
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 3), class = Intercept),
            prior(cauchy(0, 0.1), class = sd),
            prior(cauchy(0, 0.1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 6000,
  warmup = 3000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/dur_null",
  save_all_pars = TRUE
)
```

```{r dur-bf, cache=TRUE, dependson=c("dur", "dur-null")}
# calculate the Bayes factor of the difference between the full and null models
brms::bayes_factor(dur, dur_null)
```

```{r dur-post}
# calculate the marginal posteriors of the full model
dur_post <- brms::posterior_samples(dur, pars="b_") %>%
  dplyr::mutate(
    nd = exp(b_Intercept),
    nt = exp(b_Intercept) * exp(b_voicingvoiceless)
  ) %>%
  dplyr::select(nd, nt) %>%
  tidyr::gather(context, DV)
```

## Checks

```{r dur-pp}
pp_check(dur, nsamples = 50) + theme_minimal()
```

Sensitivity analysis from @Betancourt2018.
$s$ (posterior shrinkage) indicates the contribution of the data to the posterior (values closer to 1 are better).
$z$ (posterior z-score) indicates how close the inference is to the data generating process (lower values are better).
An ideal fit should have $s$ = 1 and $z$ = 0.
Posteriors with high $s$ and high $z$ indicate overfit (the data dominate the posterior), while posteriors with low $s$ and low $z$ indicate that the model is poorly identified (the data is not contributing to the model and the priors dominate).

```{r dur-sens}
dur_fixed <- fixef(dur) %>% as_tibble(rownames = "term")

dur_fixed %>%
  mutate(
    theta = c(0, 0),
    sigma_prior = c(3, 1),
    # it's called here std.error but is the standard deviation
    z = abs((Estimate - theta) / Est.Error),
    s = 1 - (Est.Error^2 / sigma_prior^2)
  ) %>%
  ggplot(aes(s, z, label = term)) +
  geom_point() +
  geom_label(nudge_x = -0.1, size = 2, alpha = 0.5) +
  xlim(0, 1) + ylim(0, 250) +
  theme_minimal()
```

# Velum gesture onset

See velum gesture duration for prior specification.

```{r onset, cache=TRUE}
# create the dependent variable
subdat$DV <- -(subdat$velumopening_gesture_on - subdat$Vokal_off)*1000

# full model
onset <- brms::brm(
  DV ~ voicing +
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 3), class = Intercept),
            prior(normal(0, 1), class = b, coef = voicingvoiceless),
            prior(cauchy(0, 0.1), class = sd),
            prior(cauchy(0, 0.1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 4000,
  warmup = 2000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/onset",
  save_all_pars = TRUE
)
```

```{r onset-summary}
onset
```

```{r onset-null, cache=TRUE}
# reduced/null model
onset_null <- brms::brm(
  DV ~
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 3), class = Intercept),
            prior(cauchy(0, 0.1), class = sd),
            prior(cauchy(0, 0.1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 4000,
  warmup = 2000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/onset_null",
  save_all_pars = TRUE
)
```

```{r onset-bf, cache=TRUE, dependson=c("onset", "onset-null")}
# calculate the Bayes factor of the difference between the full and null models
brms::bayes_factor(onset, onset_null)
```

```{r onset-post}
# calculate the marginal posteriors of the full model
onset_post <- brms::posterior_samples(onset, pars="b_") %>%
  dplyr::mutate(
    nd = exp(b_Intercept),
    nt = exp(b_Intercept) * exp(b_voicingvoiceless)
  ) %>%
  dplyr::select(nd, nt) %>%
  tidyr::gather(context, DV)
```

## Checks

```{r onset-pp}
pp_check(onset, nsamples = 50) + theme_minimal()
```

```{r onset-sens}
onset_fixed <- fixef(onset) %>% as_tibble(rownames = "term")

onset_fixed %>%
  mutate(
    theta = c(0, 0),
    sigma_prior = c(3, 1),
    # it's called here std.error but is the standard deviation
    z = abs((Estimate - theta) / Est.Error),
    s = 1 - (Est.Error^2 / sigma_prior^2)
  ) %>%
  ggplot(aes(s, z, label = term)) +
  geom_point() +
  geom_label(nudge_x = -0.1, size = 2, alpha = 0.5) +
  xlim(0, 1) + ylim(0, 250) +
  theme_minimal()
```

# Velum gesture peak (timing)

The BRM for the time point of the velum gesture peak was built using a Gaussian distribution; unlike the other measures of timing, the time point of the velum gesture peak is not expected to follow a one-sided (positive support) distribution, since the peak can potentially occur before or after the vowel offset.
The following distributions were used as weakly informative priors (on the milliseconds scale): for the intercept (corresponding to ND), $Normal(0, 200)$; for the effect of voicing (when NT), $Normal(0, 100)$.
These correspond to a belief that the intercept is between -400 and 400 ms from the vowel offset, and that the time changes by -200 to 200 ms in the NT context, at 95% confidence.
For the model standard deviation and the random intercept standard deviation we used $HalfCauchy(0, 5)$, which corresponds to a 95% HDI = [0, 127] ms.

```{r half-cauchy-get-max}
inverseCDF(c(0.025, 0.975), phcauchy, sigma = 5)
```

For the correlation between random effects, an $LKJ(2)$ distribution.

```{r gest-max, cache=TRUE}
# create the dependent variable
subdat$DV <- (subdat$velumopening_maxcon_on - subdat$Vokal_off)*1000

# full model
gest.max <- brms::brm(
  DV ~ voicing +
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = gaussian(),
  prior = c(prior(normal(0, 200), class = Intercept),
            prior(normal(0, 100), class = b, coef = voicingvoiceless),
            prior(cauchy(0, 5), class = sd),
            prior(cauchy(0, 5), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 6000,
  warmup = 3000,
  chains = 4, 
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/gest_max",
  save_all_pars = TRUE
)
```

```{r gest-max-summary}
gest.max
```

```{r gest-max-null, cache=TRUE}
# reduced/null model
gest.max_null <- brms::brm(
  DV ~
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = gaussian(),
  prior = c(prior(normal(0, 200), class = Intercept),
            prior(cauchy(0, 5), class = sd),
            prior(cauchy(0, 5), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 6000,
  warmup = 3000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/gest_max_null",
  save_all_pars = TRUE
)
```

```{r gest-max-bf, cache=TRUE, dependson=c("gest-max", "gest-max-null")}
# calculate the Bayes factor of the difference between the full and null models
brms::bayes_factor(gest.max, gest.max_null)
```

```{r gest-max-post}
# calculate the marginal posteriors of the full model
gest.max_post <- brms::posterior_samples(gest.max, pars="b_") %>%
  dplyr::mutate(
    nd = b_Intercept,
    nt = b_Intercept + b_voicingvoiceless
  ) %>%    
  dplyr::select(nd, nt) %>%
  tidyr::gather(context, DV)
```

## Checks

```{r gest-max-pp}
pp_check(gest.max, nsamples = 50) + theme_minimal()
```

```{r gest-max-sens}
gest.max_fixed <- fixef(gest.max) %>% as_tibble(rownames = "term")

gest.max_fixed %>%
  mutate(
    theta = c(0, 0),
    sigma_prior = c(200, 100),
    # it's called here std.error but is the standard deviation
    z = abs((Estimate - theta) / Est.Error),
    s = 1 - (Est.Error^2 / sigma_prior^2)
  ) %>%
  ggplot(aes(s, z, label = term)) +
  geom_point() +
  geom_label(nudge_x = -0.1, size = 2, alpha = 0.5) +
  xlim(0, 1) + ylim(0, 250) +
  theme_minimal()
```

# Velum gesture offset

See velum gesture duration for prior specification.

```{r offset, cache=TRUE}
# create the dependent variable
subdat$DV <- (subdat$velumopening_gesture_off - subdat$Vokal_off)*1000

# full model
offset <- brms::brm(
  DV ~ voicing +
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 3), class = Intercept),
            prior(normal(0, 1), class = b, coef = voicingvoiceless),
            prior(cauchy(0, 0.1), class = sd),
            prior(cauchy(0, 0.1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 6000,
  warmup = 3000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/offset",
  save_all_pars = TRUE
)
```

```{r offset-summary}
offset
```

```{r offset-null, cache=TRUE}
# reduced/null model
offset_null <- brms::brm(
  DV ~
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 3), class = Intercept),
            prior(cauchy(0, 0.1), class = sd),
            prior(cauchy(0, 0.1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 6000,
  warmup = 3000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/offset_null",
  save_all_pars = TRUE
)
```

```{r offset-bf, cache=TRUE, dependson=c("offset", "offset-null")}
# calculate the Bayes factor of the difference between the full and null models
brms::bayes_factor(offset, offset_null)
```

```{r offset-post}
# calculate the marginal posteriors of the full model
offset_post <- brms::posterior_samples(offset, pars="b_") %>%
  dplyr::mutate(
    nd = exp(b_Intercept),
    nt = exp(b_Intercept) * exp(b_voicingvoiceless)
  ) %>%
  dplyr::select(nd, nt) %>%
  tidyr::gather(context, DV)
```

## Checks

```{r offset-pp}
pp_check(offset, nsamples = 50) + theme_minimal()
```

```{r offset-sens}
offset_fixed <- fixef(offset) %>% as_tibble(rownames = "term")

offset_fixed %>%
  mutate(
    theta = c(0, 0),
    sigma_prior = c(3, 1),
    # it's called here std.error but is the standard deviation
    z = abs((Estimate - theta) / Est.Error),
    s = 1 - (Est.Error^2 / sigma_prior^2)
  ) %>%
  ggplot(aes(s, z, label = term)) +
  geom_point() +
  geom_label(nudge_x = -0.1, size = 2, alpha = 0.5) +
  xlim(0, 1) + ylim(0, 250) +
  theme_minimal()
```

# Velum gesture peak (magnitude)

The BRM for the velum gesture magnitude was built using a Beta distribution, since the magnitude values are on a 0-1 scale.
The following practically flat priors were used: $Normal(0, 10)$ for the intercept; $Normal(0, 5)$ for voicing and the random effects standard deviations; the brms default prior for the $\phi$ parameter of the beta distribution ($gamma(0.01, 0.01)$); and $LKJ(2)$ prior for the random effects correlation.

```{r gest-max-mag, cache=TRUE}
# create the dependent variable
subdat$DV <- subdat$velum2US_velumopening_maxcon_onset

# full model
gest.max.mag <- brms::brm(
  DV ~ voicing +
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = Beta(),
  prior = c(prior(normal(0, 10), class = Intercept),
            prior(normal(0, 5), class = b, coef = voicingvoiceless),
            prior(cauchy(0, 5), class = sd),
            prior(gamma(0.01, 0.01), class = phi),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 4000,
  warmup = 2000,
  chains = 4, 
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/gest_max_mag",
  save_all_pars = TRUE
)
```

```{r gest-max-mag-summary}
gest.max.mag
```

```{r gest-max-mag-null, cache=TRUE}
# reduced/null model
gest.max.mag_null <- brms::brm(
  DV ~
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = Beta(),
  prior = c(prior(normal(0, 10), class = Intercept),
            prior(cauchy(0, 5), class = sd),
            prior(gamma(0.01, 0.01), class = phi),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 4000,
  warmup = 2000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/gest_max_mag_null",
  save_all_pars = TRUE
)
```

```{r gest-max-mag-bf, cache=TRUE, dependson=c("gest-max-mag", "gest-max-mag-null")}
# calculate the Bayes factor of the difference between the full and null models
brms::bayes_factor(gest.max.mag, gest.max.mag_null)
```

```{r gest-max-mag-post}
# calculate the marginal posteriors of the full model
gest.max.mag_post <- brms::posterior_samples(gest.max.mag, pars="b_") %>%
  dplyr::mutate(
    nd = plogis(b_Intercept),
    nt = plogis(b_Intercept + b_voicingvoiceless)
  ) %>%
  dplyr::select(nd, nt) %>%
  tidyr::gather(context, DV)
```

## Checks

```{r gest-max-mag-pp}
pp_check(gest.max.mag, nsamples = 50) + theme_minimal()
```

```{r gest-max-mag-sens}
gest.max.mag_fixed <- fixef(gest.max.mag) %>% as_tibble(rownames = "term")

gest.max.mag_fixed %>%
  mutate(
    theta = c(0, 0),
    sigma_prior = c(10, 5),
    # it's called here std.error but is the standard deviation
    z = abs((Estimate - theta) / Est.Error),
    s = 1 - (Est.Error^2 / sigma_prior^2)
  ) %>%
  ggplot(aes(s, z, label = term)) +
  geom_point() +
  geom_label(nudge_x = -0.1, size = 2, alpha = 0.5) +
  xlim(0, 1) + ylim(0, 250) +
  theme_minimal()
```

# Gesture offset stiffness

```{r stiff-ons, cache=TRUE}
# create the dependent variable
subdat$DV <- subdat$stiff.ons

# full model
stiff.ons <- brms::brm(
  DV ~ voicing +
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 3), class = Intercept),
            prior(normal(0, 1), class = b, coef = voicingvoiceless),
            prior(cauchy(0, 0.1), class = sd),
            prior(cauchy(0, 0.1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 4000,
  warmup = 2000,
  chains = 4, 
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/stiff_ons",
  save_all_pars = TRUE
)
```

```{r stiff-ons-summary}
stiff.ons
```

```{r stiff-ons-null, cache=TRUE}
# reduced/null model
stiff.ons_null <- brms::brm(
  DV ~
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 3), class = Intercept),
            prior(cauchy(0, 0.1), class = sd),
            prior(cauchy(0, 0.1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 4000,
  warmup = 2000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/stiff_ons_null",
  save_all_pars = TRUE
)
```

```{r stiff-ons-bf, cache=TRUE, dependson=c("stiff-ons", "stiff-ons-null")}
# calculate the Bayes factor of the difference between the full and null models
brms::bayes_factor(stiff.ons, stiff.ons_null)
```

```{r stiff-ons-post}
# calculate the marginal posteriors of the full model
stiff.ons_post <- brms::posterior_samples(stiff.ons, pars="b_") %>%
  dplyr::mutate(
    nd = exp(b_Intercept),
    nt = exp(b_Intercept + b_voicingvoiceless)
  ) %>%
  dplyr::select(nd, nt) %>%
  tidyr::gather(context, DV)
```

## Checks

```{r stiff-ons-pp}
pp_check(stiff.ons, nsamples = 50) + theme_minimal()
```

```{r stiff-ons-sens}
stiff.ons_fixed <- fixef(stiff.ons) %>% as_tibble(rownames = "term")

stiff.ons_fixed %>%
  mutate(
    theta = c(0, 0),
    sigma_prior = c(3, 1),
    # it's called here std.error but is the standard deviation
    z = abs((Estimate - theta) / Est.Error),
    s = 1 - (Est.Error^2 / sigma_prior^2)
  ) %>%
  ggplot(aes(s, z, label = term)) +
  geom_point() +
  geom_label(nudge_x = -0.1, size = 2, alpha = 0.5) +
  xlim(0, 1) + ylim(0, 250) +
  theme_minimal()
```

# Gesture offset stiffness

```{r stiff-off, cache=TRUE}
# create the dependent variable
subdat$DV <- subdat$stiff.off

# full model
stiff.off <- brms::brm(
  DV ~ voicing +
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 3), class = Intercept),
            prior(normal(0, 1), class = b, coef = voicingvoiceless),
            prior(cauchy(0, 0.1), class = sd),
            prior(cauchy(0, 0.1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 4000,
  warmup = 2000,
  chains = 4, 
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/stiff_off",
  save_all_pars = TRUE
)
```

```{r stiff-off-summary}
stiff.off
```

```{r stiff-off-null, cache=TRUE}
# reduced/null model
stiff.off_null <- brms::brm(
  DV ~
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 3), class = Intercept),
            prior(cauchy(0, 0.1), class = sd),
            prior(cauchy(0, 0.1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 6000,
  warmup = 3000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/stiff_off_null",
  save_all_pars = TRUE
)
```

```{r stiff-off-bf, cache=TRUE, dependson=c("stiff-off", "stiff-off-null")}
# calculate the Bayes factor of the difference between the full and null models
brms::bayes_factor(stiff.off, stiff.off_null)
```

```{r stiff-off-post}
# calculate the marginal posteriors of the full model
stiff.off_post <- brms::posterior_samples(stiff.off, pars="b_") %>%
  dplyr::mutate(
    nd = exp(b_Intercept),
    nt = exp(b_Intercept) * exp(b_voicingvoiceless)
  ) %>%
  dplyr::select(nd, nt) %>%
  tidyr::gather(context, DV)
```

## Checks

```{r stiff-off-pp}
pp_check(stiff.off, nsamples = 50) + theme_minimal()
```

```{r stiff-off-sens}
stiff.off_fixed <- fixef(stiff.off) %>% as_tibble(rownames = "term")

stiff.off_fixed %>%
  mutate(
    theta = c(0, 0),
    sigma_prior = c(3, 1),
    # it's called here std.error but is the standard deviation
    z = abs((Estimate - theta) / Est.Error),
    s = 1 - (Est.Error^2 / sigma_prior^2)
  ) %>%
  ggplot(aes(s, z, label = term)) +
  geom_point() +
  geom_label(nudge_x = -0.1, size = 2, alpha = 0.5) +
  xlim(0, 1) + ylim(0, 250) +
  theme_minimal()
```

# Kurtosis

```{r kurt, cache=TRUE}
# create the dependent variable
subdat$DV <- subdat$kurtosis

# full model
kurt <- brms::brm(
  DV ~ voicing +
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 3), class = Intercept),
            prior(normal(0, 3), class = b, coef = voicingvoiceless),
            prior(cauchy(0, 1), class = sd),
            prior(cauchy(0, 1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 4000,
  warmup = 2000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/kurtosis",
  save_all_pars = TRUE
)
```

```{r kurt-summary}
kurt
```

```{r kurt-null, cache=TRUE}
# reduced/null model
kurt_null <- brms::brm(
  DV ~
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 3), class = Intercept),
            prior(cauchy(0, 1), class = sd),
            prior(cauchy(0, 1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 4000,
  warmup = 2000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/kurtosis_null",
  save_all_pars = TRUE
)
```

```{r kurt-bf, cache=TRUE, dependson=c("kurt", "kurt-null")}
# calculate the Bayes factor of the difference between the full and null models
brms::bayes_factor(kurt, kurt_null)
```

```{r kurt-post}
# calculate the marginal posteriors of the full model
kurt_post <- brms::posterior_samples(kurt, pars="b_") %>%
  dplyr::mutate(
    nd = exp(b_Intercept),
    nt = exp(b_Intercept + b_voicingvoiceless)
  ) %>%
  dplyr::select(nd, nt) %>%
  tidyr::gather(context, DV)
```

## Checks

```{r kurt-pp}
pp_check(kurt, nsamples = 50) + theme_minimal()
```

```{r kurt-sens}
kurt_fixed <- fixef(kurt) %>% as_tibble(rownames = "term")

kurt_fixed %>%
  mutate(
    theta = c(0, 0),
    sigma_prior = c(3, 1),
    # it's called here std.error but is the standard deviation
    z = abs((Estimate - theta) / Est.Error),
    s = 1 - (Est.Error^2 / sigma_prior^2)
  ) %>%
  ggplot(aes(s, z, label = term)) +
  geom_point() +
  geom_label(nudge_x = -0.1, size = 2, alpha = 0.5) +
  xlim(0, 1) + ylim(0, 250) +
  theme_minimal()
```

# Crest factor

```{r crest-fact, cache=TRUE}
# create the dependent variable
subdat$DV <- subdat$crest.fact

# full model
crest.fact <- brms::brm(
  DV ~ voicing +
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 3), class = Intercept),
            prior(normal(0, 3), class = b, coef = voicingvoiceless),
            prior(cauchy(0, 1), class = sd),
            prior(cauchy(0, 1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 4000,
  warmup = 2000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/crest_factor",
  save_all_pars = TRUE
)
```

```{r crest-fact-summary}
crest.fact
```

```{r crest-fact-null, cache=TRUE}
# reduced/null model
crest.fact_null <- brms::brm(
  DV ~
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 3), class = Intercept),
            prior(cauchy(0, 1), class = sd),
            prior(cauchy(0, 1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 4000,
  warmup = 2000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/crest_factor_null",
  save_all_pars = TRUE
)
```

```{r creast-fact-bf, cache=TRUE, dependson=c("crest-fact", "crest-fact-null")}
# calculate the Bayes factor of the difference between the full and null models
brms::bayes_factor(crest.fact, crest.fact_null)
```

```{r crest-fact-post}
# calculate the marginal posteriors of the full model
crest.fact_post <- brms::posterior_samples(crest.fact, pars="b_") %>%
  dplyr::mutate(
    nd = exp(b_Intercept),
    nt = exp(b_Intercept + b_voicingvoiceless)
  ) %>%
  dplyr::select(nd, nt) %>%
  tidyr::gather(context, DV)
```

## Checks

```{r crest-fact-pp}
pp_check(crest.fact, nsamples = 50) + theme_minimal()
```

```{r crest-fact-sens}
crest.fact_fixed <- fixef(crest.fact) %>% as_tibble(rownames = "term")

crest.fact_fixed %>%
  mutate(
    theta = c(0, 0),
    sigma_prior = c(3, 1),
    # it's called here std.error but is the standard deviation
    z = abs((Estimate - theta) / Est.Error),
    s = 1 - (Est.Error^2 / sigma_prior^2)
  ) %>%
  ggplot(aes(s, z, label = term)) +
  geom_point() +
  geom_label(nudge_x = -0.1, size = 2, alpha = 0.5) +
  xlim(0, 1) + ylim(0, 250) +
  theme_minimal()
```

# Integral of velum movement in vowel (area under the curve)

```{r integ, cache=TRUE}
# create the dependent variable
subdat$DV <- subdat$vowel.integ

# full model
integ <- brms::brm(
  DV ~ voicing +
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 5), class = Intercept),
            prior(normal(0, 5), class = b, coef = voicingvoiceless),
            prior(cauchy(0, 1), class = sd),
            prior(cauchy(0, 1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 6000,
  warmup = 3000,
  chains = 4, 
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/vowel_integ",
  save_all_pars = TRUE
)
```

```{r integ-summary}
integ
```

```{r integ-null, cache=TRUE}
# reduced/null model
integ_null <- brms::brm(
  DV ~
    (1 + voicing | speaker) +
    (1 + voicing | word),
  data = subdat,
  family = lognormal(),
  prior = c(prior(normal(0, 5), class = Intercept),
            prior(cauchy(0, 1), class = sd),
            prior(cauchy(0, 1), class = sigma),
            prior(lkj(2), class = cor)),
  seed = my.seed,
  iter = 6000,
  warmup = 3000,
  chains = 4,
  cores = core.num,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 20),
  file = "./rtMRI-velum/models/vowel_integ_null",
  save_all_pars = TRUE
)
```

```{r integ-bf, cache=TRUE, dependson=c("integ", "integ-null")}
# calculate the Bayes factor of the difference between the full and null models
brms::bayes_factor(integ, integ_null)
```

```{r integ-post}
# calculate the marginal posteriors of the full model
integ_post <- brms::posterior_samples(integ, pars="b_") %>%
  dplyr::mutate(
    nd = exp(b_Intercept),
    nt = exp(b_Intercept + b_voicingvoiceless)
  ) %>%
  dplyr::select(nd, nt) %>%
  tidyr::gather(context, DV)
```

## Checks

```{r integ-pp}
pp_check(integ, nsamples = 50) + theme_minimal()
```

```{r integ-sens}
integ_fixed <- fixef(integ) %>% as_tibble(rownames = "term")

integ_fixed %>%
  mutate(
    theta = c(0, 0),
    sigma_prior = c(3, 1),
    # it's called here std.error but is the standard deviation
    z = abs((Estimate - theta) / Est.Error),
    s = 1 - (Est.Error^2 / sigma_prior^2)
  ) %>%
  ggplot(aes(s, z, label = term)) +
  geom_point() +
  geom_label(nudge_x = -0.1, size = 2, alpha = 0.5) +
  xlim(0, 1) + ylim(0, 250) +
  theme_minimal()
```

# Plotting posteriors

```{r cols}
# colors to be used for plotting (suitable for B/W printing)
my.cols <- c("#2c7fb8","#7fcdbb")
```

## Figure 5: duration, onset, peak (timing), offset

```{r figure-5}
# gather posteriors
fig5_post <- bind_rows(
  dur_post %>%
    rename(`duration` = DV)
  %>% pivot_longer(`duration`, names_to = "outcome", values_to = "estimate"),
  onset_post %>%
    rename(`onset` = DV) %>%
    pivot_longer(`onset`, names_to = "outcome", values_to = "estimate"),
  gest.max_post %>%
    rename(`peak` = DV) %>%
    pivot_longer(`peak`, names_to = "outcome", values_to = "estimate"),
  offset_post %>%
    rename(`offset` = DV) %>%
    pivot_longer(`offset`, names_to = "outcome", values_to = "estimate")
) %>%
  mutate(outcome = factor(outcome, levels = c("offset","peak","onset","duration")))

# flip sign of velum gesture onset posteriors for interpretability
fig5_post$estimate[fig5_post$outcome=="onset"] <- -fig5_post$estimate[fig5_post$outcome=="onset"]

# make separate plots
p1 <- fig5_post %>%
  filter(outcome == "duration") %>%
  ggplot(aes(estimate, outcome, group = context, fill = context)) +
  geom_halfeyeh(slab_color="black", slab_alpha=0.5) +
  scale_x_continuous(breaks=seq(0,500,5)) +
  coord_cartesian(ylim = c(1.3, 1.5), xlim = c(250,327)) +
  scale_fill_manual(values = my.cols) +
  labs(x = "Duration (ms)", y = element_blank()) + theme_bw() +
  theme(legend.position = "none",
        axis.text=element_text(size=12),
        axis.title=element_text(size=13))

p2 <- fig5_post %>%
  filter(outcome %in% c("onset","peak","offset")) %>%
  ggplot(aes(estimate, outcome, group = context, fill = context)) +
  geom_halfeyeh(slab_color="black", slab_alpha=0.5) +
  geom_vline(xintercept = 0, lty=2) +
  scale_x_continuous(breaks=seq(-200,300,20)) +
  coord_cartesian(xlim = c(-125,205)) +
  scale_fill_manual(values=my.cols) +
  labs(x = "Time (ms) relative to acoustic vowel offset",
       y = element_blank(), fill = "Context") + theme_bw() + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=13),
        panel.grid.major.y=element_blank())
```

```{r figure-5-save}
# save coposite plot
pdf(file="./rtMRI-velum/plots/time_plots.pdf",width=9.5,height=5,
    onefile=T,pointsize=14)
(p1 + p2) + patchwork::plot_layout(ncol = 1, guides = "collect") +
  theme(legend.position = "right")
dev.off()
```

## Figure 6: peak (magnitude)

```{r figure-6}
# gather posteriors
fig6_post <- bind_rows(
  gest.max.mag_post %>%
    rename(`peak    \nmagnitude` = DV) %>%
    pivot_longer(`peak    \nmagnitude`, names_to = "outcome",
                 values_to = "estimate")
) 

# make plot
p1 <- fig6_post %>%
  filter(outcome == "peak    \nmagnitude") %>%
  ggplot(aes(estimate, outcome, group = context, fill = context)) +
  geom_halfeyeh(slab_color="black", slab_alpha=0.5) +
  scale_x_continuous(breaks=seq(0,1,0.025)) +
  coord_cartesian(xlim = c(0.58,0.775), ylim = c(1.4,1.4)) +
  scale_fill_manual(values=my.cols) +
  labs(x = "Velum opening magnitude (speaker-scaled)", y = element_blank(),
       fill = "Context") + theme_bw() +
  theme(axis.text=element_text(size=12),axis.title=element_text(size=13))
```

```{r figure-6-save}
# save plot
pdf(file="./rtMRI-velum/plots/magnitude_plot.pdf",width=9,height=3,onefile=T,
    pointsize=14)
p1
dev.off()
```

## Figure 7: opening stiffness, closing stiffness

```{r figure-7}
# gather posteriors
fig7_post <- bind_rows(
  stiff.ons_post %>% rename(`opening\nstiffness` = DV) %>%
    pivot_longer(`opening\nstiffness`, names_to = "outcome", values_to = "estimate"),
  stiff.off_post %>% rename(`closing \nstiffness` = DV) %>%
    pivot_longer(`closing \nstiffness`, names_to = "outcome", values_to = "estimate")
) 

# make composite plot on same scale
p1 <- fig7_post %>%
  filter(outcome %in% c("opening\nstiffness","closing \nstiffness")) %>%
  ggplot(aes(estimate, outcome, group = context, fill = context)) +
  geom_halfeyeh(slab_color="black", slab_alpha=0.5) +
  scale_x_continuous(breaks=seq(0,30,0.5)) +
  coord_cartesian(xlim = c(12,17.5), ylim = c(1.4,2.3)) +
  scale_fill_manual(values=my.cols) +
  labs(x = "Stiffness", y = element_blank(), fill = "Context") + theme_bw() +
  theme(axis.text=element_text(size=12),axis.title=element_text(size=13),
        panel.grid.major.y=element_blank())
```

```{r figure-7-save}
# save plot
pdf(file="./rtMRI-velum/plots/stiffness_plots.pdf",width=9,height=4,onefile=T,
    pointsize=14)
p1
dev.off()
```

## Figure 9: kurtosis, crest factor

```{r figure-9}
# gather posteriors
fig9_post <- bind_rows(
  kurt_post %>% rename(`kurtosis` = DV) %>%
    pivot_longer(`kurtosis`, names_to = "outcome", values_to = "estimate"),
  crest.fact_post %>% rename(`crest \nfactor` = DV) %>%
    pivot_longer(`crest \nfactor`, names_to = "outcome", values_to = "estimate")
) 

# make separate plots
p1 <- fig9_post %>%
  filter(outcome == "kurtosis") %>%
  ggplot(aes(estimate, outcome, group = context, fill = context)) +
  geom_halfeyeh(slab_color="black", slab_alpha=0.5) +
  scale_x_continuous(breaks=seq(0,5,0.05)) +
  coord_cartesian(xlim = c(2.81,3.18), ylim = c(1.4,1.4)) +
  scale_fill_manual(values=my.cols) +
  labs(x = "Kurtosis", y = element_blank(), fill = "Context") +
  theme_bw() +
  theme(legend.position = "none",axis.text=element_text(size=12),
        axis.title=element_text(size=13),panel.grid.major.y=element_blank())

p2 <- fig9_post %>%
  filter(outcome == "crest \nfactor") %>%
  ggplot(aes(estimate, outcome, group = context, fill = context)) +
  geom_halfeyeh(slab_color="black", slab_alpha=0.5) +
  scale_x_continuous(breaks=seq(0,3,0.01)) +
  coord_cartesian(xlim = c(1.805,1.885), ylim = c(1.4,1.4)) +
  scale_fill_manual(values=my.cols) +
  labs(x = "Crest factor (ratio of peak to average displacement)",
       y = element_blank(), fill = "Context") + theme_bw() +
  theme(legend.position = "none",axis.text=element_text(size=12),
        axis.title=element_text(size=13),panel.grid.major.y=element_blank())
```

```{r figure-9-save}
# save composite plot
pdf(file="./rtMRI-velum/plots/peakedness_plots.pdf",width=9,height=5,
    onefile=T,pointsize=14)
(p1 + p2) + patchwork::plot_layout(ncol = 1, guides = "collect") +
  theme(legend.position = "right")
dev.off()
```

## Figure 10: velum displacement integral

```{r figure-10}
# gather posteriors
fig10_post <- bind_rows(
  integ_post %>% rename(`gesture\nintegral` = DV) %>%
    pivot_longer(`gesture\nintegral`, names_to = "outcome",
                 values_to = "estimate")
) 

# make plot
p1 <- fig10_post %>%
  filter(outcome == "gesture\nintegral") %>%
  ggplot(aes(estimate, outcome, group = context, fill = context)) +
  geom_halfeyeh(slab_color="black", slab_alpha=0.5) +
  scale_x_continuous(breaks=seq(0,100,2)) +
  coord_cartesian(xlim = c(31,64.5), ylim = c(1.4,1.4)) +
  scale_fill_manual(values=my.cols) +
  labs(x = "Velum displacement integral: time (ms) X magnitude (normalized)",
       y = element_blank(), fill = "Context") + theme_bw() + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=13))
```

```{r figure-10-save}
# save plot
pdf(file="./rtMRI-velum/plots/integral_plot.pdf",width=9,height=3,onefile=T,
    pointsize=14)
p1
dev.off()
```

# References
